---
title: "Hands-on Exercise 1"
author: "Ye Zhili"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv",show_col_types = FALSE)
```

## Plotting a simple bar chart

```{r}
ggplot(data=exam_data,aes(x=RACE))+geom_bar()
```

## Dotplot

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL) 
```

## histogram

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram() 
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

## Kernel density estimate

```{r}
ggplot(data=exam_data,
aes(x = MATHS)) +
geom_density()
```

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

## Boxplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

Plotting boxplot with means

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun="mean",         
               colour ="red",        
               size=4)                         
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)           
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                  
  geom_point(position="jitter", 
             size = 0.5)                  
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()           
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin(fill="light blue") +
  geom_boxplot(alpha=0.5)                    
```

## Scatter plot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()           
```

Plotting scatterplot with best fit curve

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(size=0.5)          
```

## Facet

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)           
```

## -Box plot of math scores for classes grouped by gender

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= CLASS)) +
  geom_boxplot() +
  facet_grid(~ GENDER)       
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= CLASS)) +
  geom_boxplot() +
  facet_grid(GENDER ~.)       
```
