---
title: "Take-home_Ex01"
author: "Ye Zhili"
---

Data Preparation First, import relevant library and the dataset.

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(forcats)
library(readxl)
raw_data <- read_excel("Upload for elsiver.xlsx")
```

Second, we need to check if there is missing value in each column.

```{r}
na_summary <- raw_data %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "NA_Count") %>%
  mutate(NA_Percent = round(NA_Count / nrow(raw_data) * 100, 1)) %>%
  arrange(desc(NA_Count))
print(na_summary, n = 30)
```

Third,drop columns with high missing rates.Living Conditions(LHO1-4, LCRE, LSAV, LWDA) give me 22%-90% missing value so I will drop this section. Nearly 30% data are missing in PPO1 and PPO2 so I will also drop theses two variables.

```{r}
df <- raw_data %>% select(-LHO1, -LHO2, -LHO3, -LHO4, -LCRE, -LSAV, -LWDA, -PPO1, -PPO2)
```

Forth, some variables needed data type conversion for subsequent data analysis. Four variables were primarily converted:

1.  CGEN (Gender): The original data consisted of text "1" and "2" instead of numbers, so it was converted to numeric.

2.  ARO4 (Better living conditions in rural areas): A backtick caused the entire column to become character data. `na_if()` converts empty strings and backticks to NA, and `as.numeric()` converts the remaining "1", "2", and "3" from text to numbers.

3.  ARO2 (Increased Woman roles in their families): One record had a value of 11. Normally, there are only three options: 1, 2, and 3, so this was deleted as an outlier.

4.  TOIN (Other Income): Income cannot be negative, so this was also deleted as an outlier.

```{r}
df <- df %>%
  mutate(
    CGEN = as.numeric(CGEN),
    ARO4 = as.numeric(na_if(na_if(ARO4, ""), "`"))) %>%
    filter(ARO2 != 11 | is.na(ARO2)) %>%             
    filter(TOIN >= 0 | is.na(TOIN)) 
```

Fifth,remove rows with any NA

```{r}
df_clean <- df %>% drop_na()
```

Six,add meaningful labels

```{r}
df_clean <- df_clean %>%
  mutate(
    Province = factor(CPRO, levels = 1:5,
                      labels = c("Tuyen Quang", "Quang Ninh", "Ha Giang",
                                 "Yen Bai", "Bac Giang")),
    Gender = factor(CGEN, levels = 1:2,
                    labels = c("Male", "Female")),
    Race = factor(CRAC, levels = 1:2,
                  labels = c("Kinh", "Minorities")),
    Career = factor(CJOB, levels = 1:3,
                    labels = c("Agricultural", "Service & Industrial", "Others")),
    Quintile = factor(CQUI, levels = 1:5,
                      labels = c("Top", "Fourth", "Middle", "Second", "Lowest")),
    Education = factor(FEDU, levels = 0:3,
                       labels = c("Primary", "Lower Secondary",
                                  "Upper Secondary", "Others/Higher")),
    Training = factor(FVTP, levels = 1:2,
                      labels = c("Short Courses", "Long Courses")),
    Credit = factor(FCRA, levels = 1:2,
                    labels = c("Inaccessible", "Accessible")),
    Technology = factor(FTAP, levels = 1:2,
                        labels = c("Low", "Adequate")),
    Income_Group = factor(
      case_when(CQUI %in% c(1, 2) ~ "High Income",
                CQUI == 3          ~ "Middle Income",
                CQUI %in% c(4, 5)  ~ "Low Income"),
      levels = c("High Income", "Middle Income", "Low Income"))
  )
```

Data Analysis and Visualization Research Question 1:Do different provinces differ in ethnic composition & career structure?

::: panel-tabset
## Ethnicity by Province

```{r}
#| code-fold: true
#| code-summary: "Show the code"
eth_data <- df_clean %>%
  count(Province, Race) %>%
  group_by(Province) %>%
  mutate(Total = sum(n), Pct = n / Total * 100,
         Label = paste0(n, "\n(", round(Pct, 1), "%)"))

ggplot(eth_data, aes(x = Province, y = Pct, fill = Race)) +
  geom_col(position = "stack", width = 0.65) +
  geom_text(aes(label = Label), position = position_stack(vjust = 0.5),
            size = 3.2, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Kinh" = "#2E86AB", "Minorities" = "#E8505B")) +
  labs(title = "Ethnic Composition by Province",
       subtitle = "Percentage of Kinh vs Ethnic Minorities among surveyed labourers",
       x = NULL, y = "Percentage (%)", fill = "Ethnicity") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

## Career by Province
```{r}
#| code-fold: true
#| code-summary: "Show the code"
career_data <- df_clean %>%
  count(Province, Career) %>%
  group_by(Province) %>%
  mutate(Total = sum(n), Pct = n / Total * 100,
         Label = ifelse(Pct >= 8, paste0(n, "\n(", round(Pct, 1), "%)"), ""))

ggplot(career_data, aes(x = Province, y = Pct, fill = Career)) +
  geom_col(position = "stack", width = 0.65) +
  geom_text(aes(label = Label), position = position_stack(vjust = 0.5),
            size = 3.2, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Agricultural" = "#4CAF50",
                                "Service & Industrial" = "#FF9800",
                                "Others" = "#9C27B0")) +
  labs(title = "Career Structure by Province",
       subtitle = "Distribution of occupational sectors among surveyed labourers",
       x = NULL, y = "Percentage (%)", fill = "Career") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

## Ethnicity × Career
```{r}
#| code-fold: true
#| code-summary: "Show the code"

cross_data <- df_clean %>%
  count(Race, Career) %>%
  group_by(Race) %>%
  mutate(Total = sum(n), Pct = n / Total * 100,
         Label = paste0(n, " (", round(Pct, 1), "%)"))

ggplot(cross_data, aes(x = Race, y = Pct, fill = Career)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = Label), position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3.2) +
  scale_fill_manual(values = c("Agricultural" = "#4CAF50",
                                "Service & Industrial" = "#FF9800",
                                "Others" = "#9C27B0")) +
  labs(title = "Career Structure: Kinh vs Ethnic Minorities",
       subtitle = "Do ethnic minorities have less access to non-agricultural employment?",
       x = NULL, y = "Percentage (%)", fill = "Career") +
  ylim(0, 100) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```
:::
